A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-18 paper/paper.Rmd
tidied up the new section on hgc

`new_data_qnames`, contains the variables `Q1-3_A~Y_1979` and `Q1-3_A~Y_1981` which records two versions of the birth year of the respondent; this is also the case for the record of birth month (`Q1-3_A~M_1979` and `Q1-3_A~M_1981`). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

@@

The refreshed dataset has three types of highest grade completed. The first one is the highest grade completed ever reported in the database (`hgc` and `hgc_i`, for the factor and integer type, respectively). For each ID, there is only one value of `hgc` and `hgc_i`. This variable is obtained from `new_data_qnames` with name `HGC_EVER_XRND` and stored in year unit (e.g., 10, 11, 12, 13, and so on), we only rename the column name to be `hgc_i` and recode it so we have the factor data type (e.g., 10th grade, 11th grade, and so on).

The second one is the highest grade completed in 1979 (`hgc_1979`), the first round of the survey. We decide to include this variable in the dataset for comparison with the original data-purpose. The original data only covers the period until 1994. Hence, `hgc` in the refreshed data, which is the state of `hgc` in 2018, and the `hgc` in the original data will not be comparable.

The third one is the highest grade completed corresponds to value  reported in each round of the survey (`grade`), meaning that the value can change. We decide to include this variable in the refreshed dataset to enrich the analysis possibly done with the data, for example if one would like to explore how the changes of highest grade completed in individuals affecting their wages. The highest grade completed are recorded in `new_data_qnames` as variables beginning with `Q3-4` and `HGC` with suffix of the year it was recorded. In addition the variables beginning with `HGCREV` contain the revised data. We choose to use this revised data because it seemed to have less missing and presumably has been checked. However, there is no revised data for 2012, 2014, 2016, and 2018, thus, we use the ordinary data for these years. <!--We also check the consistency between `hgc_i` and `grade`. Some problematic observations were found, for example, `grade` value is greater than `hgc_i`. We then fix this problem by replacing those `grade` that is greater than its `hgc_i`. -->

@@

 <!--AE comment: Not panel 9 --> the IDs in panel 9. 
<!--I think there is a narrative step missing here, which is that looking at those samples led you to look at plots that show values for the wage variable over all participants and all years. The phrase “the summary plots” was not enough description for me to know what to expect when I flipped to Figure 3. Please be explicit about what the three plots were, why they were made, and what you were looking for. I think there is another narrative step missing about Figure 3C, which is that after you saw there were outliers you tried to identify which respondents had the extreme values. Again, please spell this out in the text. When you return to the plots for the entire dataset in Figure 5, please make parallel comments. -->

>>>>

The object `new_data_qnames`, contains the variables `Q1-3_A~Y_1979` and `Q1-3_A~Y_1981` which records two versions of the birth year of the respondent; this is also the case for the record of birth month (`Q1-3_A~M_1979` and `Q1-3_A~M_1981`). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

@@

The next step is processing highest grade completed. There are several ways to define this, and this shoudl be reflected in the refreshed data, to giev some flexibility for downstream analysis. The first one is the highest grade ever completed is reported in the database and provided in the refreshed data as `hgc` and `hgc_i`, for the factor and integer type, respectively. For each individual, there is only one value of `hgc` and `hgc_i`. This variable is obtained from `new_data_qnames` with name `HGC_EVER_XRND` and stored in year units (e.g., 10, 11, 12, 13, and so on), so the transformation is simply giving a new column name as `hgc_i` and recoded as a factor to give `hgc` (e.g., 10th grade, 11th grade, and so on).

The second definition is the highest grade completed in 1979 (provided in the refreshed data as `hgc_1979`), corresponding to the value from the first round of the survey. This best reflects the grade when the individual left school, and is included in order to compare with the original data.

The third definition is the highest grade completed that is reported in each round of the survey (provided in the refreshed data as `grade`). This value can change over time, but it should only increase. This has been included in the refreshed dataset to enable richer downstream analyses, for example if one would like to explore how the temporal changes in education level affects the wages of individuals. This is recorded in `new_data_qnames` as columns beginning with `Q3-4` and `HGC` with year as a suffix. In addition, it is also in the columns beginning with `HGCREV` reflecting revised data. We choose to use this revised values because there were less missing values indicating it had been more thoroughly checked. When this value was missing, 2012, 2014, 2016, and 2018, but available in the first form substituted accordingly. <!--We also check the consistency between `hgc_i` and `grade`. Some problematic observations were found, for example, `grade` value is greater than `hgc_i`. We then fix this problem by replacing those `grade` that is greater than its `hgc_i`. -->

@@

 <!--AE comment: Not panel 9  the IDs in panel 9. -->
<!--I think there is a narrative step missing here, which is that looking at those samples led you to look at plots that show values for the wage variable over all participants and all years. The phrase “the summary plots” was not enough description for me to know what to expect when I flipped to Figure 3. Please be explicit about what the three plots were, why they were made, and what you were looking for. I think there is another narrative step missing about Figure 3C, which is that after you saw there were outliers you tried to identify which respondents had the extreme values. Again, please spell this out in the text. When you return to the plots for the entire dataset in Figure 5, please make parallel comments. -->

### 2022-02-18 paper/paper.tex
tidied up the new section on hgc

In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age\_1979}), highest grade completed reported in each round of survey (\texttt{grade}), highest grade completed ever reported (\texttt{hgc}), highest grade completed in terms of years, e.g.~9th grade = 9, 3rd year college = 15, (\texttt{hgc\_i}), and whether the graduate equivalency diploma is obtained (\texttt{ged}).

It is worth noting that \texttt{gender} comes from a variable called \texttt{sex} in the database. We use the term \texttt{gender} as the reports about the data on various pages in the website more commnly use \texttt{gender}. Gender, as provided in the data, is self-reported and only has two categories, ``male'' and ``female''. The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification (Kennedy et al. 2020). From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex (Kennedy et al. 2020). Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similar this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

The raw data, \texttt{new\_data\_qnames}, contains the variables \texttt{Q1-3\_A\textasciitilde{}Y\_1979} and \texttt{Q1-3\_A\textasciitilde{}Y\_1981} which records two versions of the birth year of the respondent; this is also the case for the record of birth month (\texttt{Q1-3\_A\textasciitilde{}M\_1979} and \texttt{Q1-3\_A\textasciitilde{}M\_1981}). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

The refreshed dataset has two types of highest grade completed. The first one is the highest grade completed ever reported in the database (\texttt{hgc} and \texttt{hgc\_i}, for the factor and integer type, respectively). For each ID, there is only one value of \texttt{hgc} and \texttt{hgc\_i}. This variable is obtained from \texttt{new\_data\_qnames} with name \texttt{HGC\_EVER\_XRND} and stored in year unit (e.g., 10, 11, 12, 13, and so on), we only rename the column name to be \texttt{hgc\_i} and recode it so we have the factor data type (e.g., 10th grade, 11th grade, and so on).

The second one is the highest grade completed corresponds to value reported in each round of the survey (\texttt{grade}), meaning that the value can change. We decide to include this variable in the refreshed dataset to enrich the analysis possibly done with the data, for example if one would like to explore how the changes of highest grade completed in individuals affecting their wages. The highest grade completed are recorded in \texttt{new\_data\_qnames} as variables beginning with \texttt{Q3-4} and \texttt{HGC} with suffix of the year it was recorded. In addition the variables beginning with \texttt{HGCREV} contain the revised data. We choose to use this revised data because it seemed to have less missing and presumably has been checked. However, there is no revised data for 2012, 2014, 2016, and 2018, thus, we use the ordinary data for these years.

@@

the IDs in panel 9.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \texttt{id}: A unique individual's ID number.
\item
  \texttt{age\_1979}: The age of the individual in 1979.
\item
  \texttt{gender}: Gender of the individual, FEMALE and MALE.
\item
  \texttt{race} : Race of the individual, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
\item
  \texttt{hgc}: The highest grade completed ever.
\item
  \texttt{hgc\_i}: Integer value of the highest grade completed ever.
\item
  \texttt{ged}: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
\end{enumerate}

>>>>

In our final target data, we wish to include the demographic variables with variable names specified in brackets: gender (\texttt{gender}), race (\texttt{race}), age (\texttt{age\_1979}), highest grade completed reported in each round of survey (\texttt{grade}), highest grade completed ever reported (\texttt{hgc}), highest grade completed in terms of years, e.g.~9th grade = 9, 3rd year college = 15, (\texttt{hgc\_i}), highest grade completed in 1979 (\texttt{hgc\_1979}) and whether the graduate equivalency diploma is obtained (\texttt{ged}).

For \texttt{gender} and \texttt{race}, we only rename the column names provided in the raw data, \texttt{new\_data\_qnames}. It is worth noting that \texttt{gender} comes from a variable called \texttt{sex} in the database. We use the term \texttt{gender} as the reports about the data on various pages in the website more commonly use \texttt{gender}. Gender, as provided in the data, is self-reported and only has two categories, ``male'' and ``female''. The interchangeable use of these terms reflects that gender and sex are sometimes regarded as similar in surveys, although they are not. Measuring gender as a binary variable has the potential to fail to capture people who do not identify themselves as either male or female or people whose gender does not align with their sex classification (Kennedy et al. 2020). From a statistical perspective, this sometimes difficult the adjustment of survey statistics to the population when gender is measured, for example, with three categories, and the census measured it with only two categories and used the term sex (Kennedy et al. 2020). Hence, for modern societal studies, it is highly recommended to separate sex and gender in the survey question and provide more options to respond to questions on gender. This discourse should also be emphasized in teaching that using these two variables should be accompanied by special attention. Similar this also relates to race, as reported in the database. When doing an analysis with this variable, one should keep in mind that the purpose is to study racism rather than race.

The object \texttt{new\_data\_qnames}, contains the variables \texttt{Q1-3\_A\textasciitilde{}Y\_1979} and \texttt{Q1-3\_A\textasciitilde{}Y\_1981} which records two versions of the birth year of the respondent; this is also the case for the record of birth month (\texttt{Q1-3\_A\textasciitilde{}M\_1979} and \texttt{Q1-3\_A\textasciitilde{}M\_1981}). The record contains two versions of birth year and birth month as the survey recorded this in 1979 and 1981. We checked for consistency between the two versions and found no discrepancy where the responses were recorded in both 1979 and 1981. The age was then calculated using the birth year.

The next step is processing highest grade completed. There are several ways to define this, and this shoudl be reflected in the refreshed data, to giev some flexibility for downstream analysis. The first one is the highest grade ever completed is reported in the database and provided in the refreshed data as \texttt{hgc} and \texttt{hgc\_i}, for the factor and integer type, respectively. For each individual, there is only one value of \texttt{hgc} and \texttt{hgc\_i}. This variable is obtained from \texttt{new\_data\_qnames} with name \texttt{HGC\_EVER\_XRND} and stored in year units (e.g., 10, 11, 12, 13, and so on), so the transformation is simply giving a new column name as \texttt{hgc\_i} and recoded as a factor to give \texttt{hgc} (e.g., 10th grade, 11th grade, and so on).

The second definition is the highest grade completed in 1979 (provided in the refreshed data as \texttt{hgc\_1979}), corresponding to the value from the first round of the survey. This best reflects the grade when the individual left school, and is included in order to compare with the original data.

The third definition is the highest grade completed that is reported in each round of the survey (provided in the refreshed data as \texttt{grade}). This value can change over time, but it should only increase. This has been included in the refreshed dataset to enable richer downstream analyses, for example if one would like to explore how the temporal changes in education level affects the wages of individuals. This is recorded in \texttt{new\_data\_qnames} as columns beginning with \texttt{Q3-4} and \texttt{HGC} with year as a suffix. In addition, it is also in the columns beginning with \texttt{HGCREV} reflecting revised data. We choose to use this revised values because there were less missing values indicating it had been more thoroughly checked. When this value was missing, 2012, 2014, 2016, and 2018, but available in the first form substituted accordingly.

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \texttt{id}: A unique individual's ID number.
\item
  \texttt{age\_1979}: The age of the individual in 1979.
\item
  \texttt{gender}: Gender of the individual, FEMALE and MALE.
\item
  \texttt{race} : Race of the individual, NON-BLACK,NON-HISPANIC; HISPANIC; BLACK.
\item
  \texttt{hgc}: The highest grade completed ever.
\item
  \texttt{hgc\_i}: Integer value of the highest grade completed ever.
\item
  \texttt{hgc\_1979}: The highest grade completed in 1979 (integer value).
\item
  \texttt{ged}: Wether the respondent had a high school diploma or Graduate Equivalency Degree (GED). 1: High school diploma; 2: GED; 3: Both.
\end{enumerate}
